Improving Software Maintainability through Automated Refactoring of Duplicate Code

### 2020-01-11 ecoop/paper.tex
Future work

We identify the factors of interest characterizing the impact on the maintainability of automatically applied refactorings. We find that the biggest influencing factor is the size of the clone. The majority of duplicates larger than 63 tokens improve maintainability when refactored. The amount of external data that needs to be passed to the extracted method is the other important factor: the majority of extracted methods requiring more than two parameters decrease maintainability.
\end{abstract}

@@

We measure several maintainability metrics before and after refactoring to determine the impact on system maintainability of applying the refactoring. These metrics are volume, method parameters, duplication and complexity. Using this setup, we gain empirical data about the maintainability improvement of clones on a large set of data in a controlled environment. Because the refactoring is done automatically, we also consider micro-clones, which are shown to be important for refactoring consideration \cite{mondai2018micro}. We use this data to define which characteristics of clones have the biggest influence of system maintainability when such a clone is refactored. This allows for more a accurate suggestion of code clones for refactoring and can assist in the refactoring process.

Most clone detection tools can be configured using thresholds. These thresholds indicate the minimum number of lines, tokens and/or statements that must be spanned for duplicate fragments to be considered clones. Often, such thresholds are intuitively chosen~\cite{li2006cp, roy2009mutation} or based on a quantile distribution of empirical data~\cite{alves2010deriving}. Using the maintainability score we can find support for which thresholds should be chosen to increase the chance to find clones that improve maintainability when refactored.% \todo{Check whether this really fits here}

Over a corpus of 2.267 Java project our tool refactored 12.710 clone classes. We found that the main influencing factor on maintainability when refactored is token volume. We define token volume as the combined number of tokens in all clone instances in a clone class. When refactoring clones with a very small token volume, we found that the maintainability decreases when such clones are refactored. The average clone with a clone volume of 65 or larger improves maintainability when refactored. The other factor with a major influence on maintainability is the number of parameters that are required for the extracted method. We found that if a clone requires more than 2 parameters in the extracted method, it is more likely to decrease maintainability.

@@

@@

\subsection{Future Work}

>>>>

We identify the factors of interest characterizing the impact on the maintainability of automatically applied refactorings. We find that the biggest influencing factor is the size of the clone. The majority of duplicates with 63 or more tokens improve maintainability when refactored. The amount of external data that needs to be passed to the extracted method is the other important factor: the majority of extracted methods requiring more than two parameters decrease maintainability.
\end{abstract}

@@

We measure several maintainability metrics \cite{heitlager2007practical} before and after refactoring each clone to determine its impact on system maintainability. These metrics are volume, number of parameters, duplication and complexity. Using this setup, we gain empirical data about the maintainability improvement of clones on a large set of projects in a controlled environment. We use this data to define which characteristics of clones have the biggest influence on system maintainability when such a clone is refactored. This allows for more a accurate suggestion of code clones for refactoring and can assist in the refactoring process.

Most clone detection tools can be configured using thresholds. These thresholds indicate the minimum number of lines, tokens and/or statements that must be spanned for duplicate fragments to be considered clones. Often, such thresholds are intuitively chosen~\cite{li2006cp, roy2009mutation} or based on a quantile distribution of empirical data~\cite{alves2010deriving}. Using the maintainability score we can find support for which thresholds should be chosen to increase the chance to find clones that improve maintainability when refactored. Because of this, we minimize limiting the size of clones we find to measure the maintainability impact of all clones, including micro-clones \cite{mondai2018micro}.

Over a corpus of 2,267 Java project our tool refactored 12,710 clone classes. We found that the main influencing factor on maintainability when refactored is token volume. We define token volume as the combined number of tokens in all clone instances in a clone class. %When refactoring clones with a very small token volume, we found that the maintainability decreases when such clones are refactored. 
The average clone with a token volume of 63 or larger improves maintainability when refactored. The other factor with a major influence on maintainability is the number of parameters that are required for the extracted method. We found that if a clone requires more than 2 parameters in the extracted method, it is more likely to decrease maintainability.

@@

Fig.~\ref{fig:cloneclasses} shows an example of two clone classes. The first clone class spans the first two lines of all three files. The second clone class spans all lines of the first two files.

\begin{figure}[H]
\begin{parcolumns}{3}
\colchunk[1]{
\begin{javacode}
// File1.java
|\highlightDarkyellow|doA();
|\highlightDarkyellow|doB();
|\highlightYellow|doC();
\end{javacode}}
\colchunk[2]{
\begin{javacode}
// File2.java
|\highlightDarkyellow|doA();
|\highlightDarkyellow|doB();
|\highlightYellow|doC();
\end{javacode}}
\colchunk[3]{
\begin{javacode}
// File3.java
|\highlightYellow|doA();
|\highlightYellow|doB();
doD();
\end{javacode}}
\end{parcolumns}
\caption{Example of two clone classes.}
\label{fig:cloneclasses}
\end{figure}

@@

\subsection{Future Work}
In this study, we limited our scope by taking the naming of generated declarations outside of the scope. When using our work for refactoring assistance, these names will have to be manually provided. Recent studies aim at solving this problem by generating a name that matches the body of a declaration \cite{allamanis2015suggesting, alon2018code2seq}. We will investigate the possibility to combine such approaches with CR, to make it more usable in the refactoring process.

We also limited our scope by choosing a set of four metrics. However, software maintainability depends on more factors and can be measured by more metrics. By taking more metrics into account, we can give a better indication whether the maintainability improves after refactoring. For instance, we think that module coupling \cite{heitlager2007practical} is an important metric to be considered when refactoring clones. In out future work we will investigate other metrics and map their impact on the results.
